### 2020-11-21 latex/acrobot.tex
Finished first edit of oscillation

Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]
%TODO: EDIT FROM HERE
Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 
\(\pi\)-periodic in \(\alpha\), so that the following properties hold:
\begin{equation}\label{eqn:bhat-integral-properties}
    \begin{split}
    \int \limits_0 ^ {N \pi} \hat{b}(\sigma,\mu_0) d\sigma
    &= N \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma \, 
    \left(\forall N \in \mathbb{Z}_{> 0}\right)
    , \\
    \int \limits_{-\pi}^0 \hat{b}(\sigma,\mu_0) d\sigma
    &= \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    .
    \end{split}
\end{equation}
Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
After half an orbit, when \(\alpha = \pi\), the acrobot will have a new
pseudo-radius \(m = \hat{\mu}(\pi,\mu_0)\).
Hence, the change in pseudo-radius is 
\(m - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
After an additional half orbit, the acrobot lands on the positive \(q_u\)-axis
with pseudo-radius
\(\tilde{m} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
The properties outlined in ~\eqref{eqn:bhat-integral-properties} imply that
\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
This means the change in pseudo-radius after one entire oscillation is 
\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0) \approx 2 (m - \mu_0)\).
Hence, if \(\hat{\mu}_1(\pi,\cdot)\) is always positive,
\(\hat{\mu}\) must be increasing every half oscillation.
This in turn would guarantee \(q_u\) is increasing, which means the acrobot
will eventually reach a rotation.

To prove this intuitive argument rigorously, we define the Poincar\'{e} section
\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
in Figure \ref{fig:acrobot-Q}. 
We see that \(Q(\mu_0)\) is strictly positive and monotonically increasing, with
an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

@@

We can therefore state that, for each \(\mu_0 \in \, ]0,\pi[\), there exists
\(I\) such that
\[
    P_\mathcal{O}(\mu_0) = \mu_0\left(1+ IKQ(\mu_0)\right) + R(\pi,\mu_0,I) > 0
    .
\]
Unfortunately, Khalil's perturbation theory does not guarantee this \(I\) is the
same for all \(\mu_0\).

We deal with this issue by taking the compact set
\[
    \bar{\mathcal{O}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(q_u,p_u) \leq E(\pi,0) \right\}
    .
\]
The remainder term \(R(\pi,\mu_0,I)\) is smooth and of order \(I^2\),
which means it can be written in the form 
\(R(\pi,\mu_0,I) = I^2\tilde{R}(\mu_0,I)\)
for some \(\tilde{R}\) a smooth function.
Since \(\bar{O}\) is compact, \(\tilde{R}\) attains its minimum
\(\underbar{r}\) on \(\bar{O}\).
Hence, \(R(\pi,\mu_0,I)\) is bounded below by the parabola
\(\underbar{r}I^2\), which means
\begin{equation}\label{eqn:acrobot-Po-lowerbound}
    P_\mathcal{O}(\mu_0) \geq \mu_0\left(1 + IKQ(\mu_0)\right) + I^2\underbar{r}
    .
\end{equation}
Because \(I\) is assumed to be small, the term \(I^2\) shrinks rapidly.
If \(\underbar{r} < 0\), we can find a value of \(I\) which overcomes
the effect of \(I^2\underbar{r}\) and make \(P_\mathcal{O}\) positive.
Furthermore, ~\eqref{eqn:acrobot-Po-lowerbound} implies that, for fixed \(I\),
\(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
any \(\epsilon > 0\), there exists \(I(\epsilon)\) small enough so that
the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
\(\mu_0 \in ]\epsilon,\pi[\) and 
\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).

@@

Since \(\bar{\mathcal{O}}_\epsilon \subset \mathcal{O}\), we still have 
\(\dot{\alpha} > 0\), which means \(x(t)\) spirals clockwise around the origin.
If this orbit remains in \(\mathcal{O}\), it must intersect
the \(q_u\) axis at some radius \(\mu = \delta\)
which depends on the initial condition \(x(0)\).
Take 
\[
    \underbar{\delta} = 
    \min\limits_{x(0) \in \bar{\mathcal{O}}_\epsilon} \delta(x(0))
    ,
\] 
which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
From the previous perturbation analysis,
there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
is positive for all \(\mu_0 \in ]\underbar{\delta},\pi[\).

Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
intersects the \(q_u\)-axis at some point \(x(t_0) = (m_0, 0)\).
The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
\(|m_0| > \underbar{\delta}\).
As the acrobot continues its oscillation, it might exit
\(\mathcal{O}\) if it has gained enough energy;
if not, it will intersect the \(q_u\) axis again at
a new point \(x(t_1) = (m_1,0)\).
Theorem \ref{thm:khalil-perturbation} guarantees that \(|m_1|\) differs
from \(P_\mathcal{O}(|m_0|)\) only up to order \(I^2\), 
which means \(|m_1| > |m_0|\).
From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
initializing the acrobot at \(x(t_i) = (m_i, 0)\), the
next intersection with the \(q_u\) axis will be at
\(x(t_{i+1}) = (m_{i+1},0)\) with \(|m_{i+1}| > |m_i|\).

@@

Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
decreasing.
Hence, the orbit in \(\mathcal{O}_\epsilon\) spirals clockwise towards the origin, 
which means the constraint is dissipating energy.

>>>>

Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to ~\eqref{eqn:acrobot-mu1-dot} is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]

Suppose we initialize the acrobot at \((\alpha,\mu) = (0,\mu_0)\). 
After half an oscillation, when \(\alpha = \pi\), the orbit will land on the
negative \(q_u\)-axis with pseudo-radius 
\(m = \hat{\mu}(\pi,\mu_0) \approx \mu_0 + I\hat{\mu}_1(\pi,\mu_0)\).
Hence, the change in pseudo-radius is 
\(m - \mu_0 \approx I\hat{\mu}_1(\pi,\mu_0)\).
If the orbit remains in \(\mathcal{O}\) for the full oscillation, 
it will land on the positive \(q_u\)-axis
with pseudo-radius
\(\tilde{m} \approx \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).

Notice that \(\hat{b}(\alpha,\mu_0)\) is even and 
\(\pi\)-periodic in \(\alpha\), so that
\begin{equation}\label{eqn:bhat-integral-properties}
    \int \limits_{-\pi}^\pi \hat{b}(\sigma,\mu_0) d\sigma
    = \int \limits_0 ^ {2 \pi} \hat{b}(\sigma,\mu_0) d\sigma
    = 2 \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma
    .
\end{equation}
This implies that 
\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\),
which means the change in pseudo-radius after a full oscillation is 
\(\tilde{m} - \mu_0 \approx 2I\hat{\mu}_1(\pi,\mu_0)\).
If \(\hat{\mu}_1(\pi,\mu_0) > 0\), 
then both \(m - \mu_0 > 0\) and \(\tilde{m} - \mu_0 > 0\).
In other words, we can study the average behaviour of \(\hat{\mu}\) by examining
the behaviour over each half oscillation.

We define the Poincar\'{e} section 
\(P_\mathcal{O}(\mu_0) := \hat{\mu}(\pi,\mu_0)\)
corresponding to the change in \(\hat{\mu}\) after half an oscillation.
The section expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters, while \(\hat{b}(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
in Figure \ref{fig:acrobot-Q}. 
We see that it is strictly positive and monotonically increasing, with
an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

@@

This fact, coupled with knowledge that \(R(\pi,\mu_0,I)\) is \(O(I^2)\), 
implies there exists \(I > 0\) small enough that
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + IKQ(\mu_0) + R(\pi,\mu_0,I) > 0
    .
\]
Unfortunately, this \(I\) may not be the
same for all \(\mu_0 \in \, ]0,\pi[\), since
the value of \(R(\pi,\mu_0,I)\) could become more negative
for other initial conditions.
We deal with this issue by taking the compact set
\[
    \bar{\mathcal{O}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(q_u,p_u) \leq E(\pi,0) \right\}
    .
\]
The remainder term \(R(\alpha,\mu_0,I)\) is smooth and of order \(I^2\),
which means it can be written in the form 
\(R(\alpha,\mu_0,I) = I^2\tilde{R}(\alpha,\mu_0,I)\)
for some \(\tilde{R}\) a smooth function.
Since \(\bar{\mathcal{O}}\) is compact, \(\tilde{R}\) attains its minimum
\({r}\) on \(\bar{\mathcal{O}}\).
Hence, \(R(\alpha,\mu_0,I)\) is bounded below on \(\mathcal{O}\) by
\(I^2{r}\), which implies 
\begin{equation}\label{eqn:acrobot-Po-lowerbound}
    P_\mathcal{O}(\mu_0) \geq \mu_0 + IKQ(\mu_0) + I^2{r}
    .
\end{equation}
This equation implies that, for fixed \(I\),
\(P_\mathcal{O}(\mu_0)\) is monotonically increasing in \(\mu_0\).
Thus, for every \(\epsilon > 0\) and \(\mu_0 \in \, [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) \geq P_\mathcal{O}(\epsilon)
    = \epsilon + IKQ(\epsilon) + I^2{r}
    .
\]
Since \(I\) is assumed to be small, \(I^2\) shrinks rapidly.
Therefore, we can always find a small enough \(I > 0\) making 
\(\epsilon + I^2{r} \geq 0\), so that
\[
    P_\mathcal{O}(\epsilon) \geq IKQ(\epsilon) > 0
    .
\]

We conclude that, for all \(m\), \(g\), \(l\), \(\bar{q}_a\), and for
any \(\epsilon > 0\), there exists \(I(\epsilon)\) small enough so that
the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is positive for all
\(\mu_0 \in [\epsilon,\pi[\) and 
\(\lim \limits_{\mu_0 \to \pi} P_\mathcal{O}(\mu_0) = \infty\).

@@

Let \(x(0) \in \mathcal{O}_\epsilon \subset \mathcal{O}\).
Since \(\dot{\alpha} > 0\), the orbit \(x(t)\) spirals clockwise around
the origin.
If this orbit remains in \(\mathcal{O}\), it must intersect
the \(q_u\) axis at some radius \(\mu = \delta\)
which depends on the initial condition \(x(0)\).
Take 
\[
    \underbar{\delta} = 
    \min\limits_{x(0) \in \bar{\mathcal{O}}_\epsilon} \delta(x(0))
    ,
\] 
which exists because \(\bar{\mathcal{O}}_\epsilon\) is compact.
From the previous perturbation analysis,
there exists \(I\) small enough that \(P_\mathcal{O}(\mu_0)\) 
is positive for all \(\mu_0 \in [\underbar{\delta},\pi[\).

Suppose \(x(0) \in \mathcal{O}_\epsilon\), and the orbit
intersects the \(q_u\)-axis at some point \(x(t_0) = (m_0, 0)\).
The Poincar\'{e} section \(P_\mathcal{O}(|m_0|)\) is positive because 
\(|m_0| \geq \underbar{\delta}\).
As the acrobot continues its oscillation, it will intersect the \(q_u\) axis
again at a new point \(x(t_1) = (m_1,0)\).

Setting \(p_1 := P_\mathcal{O}(|m_0|)\),
Theorem \ref{thm:khalil-perturbation} guarantees that
\[
    \norm{m_1 - p_1} \leq cI^2
    ,
\]
for some \(c > 0\).
%TODO 
\textbf{SHOW THIS IMPLIES |m1| > |m0|}

From here, we can apply Theorem \ref{thm:khalil-perturbation} iteratively:
initializing the acrobot at \(x(t_i) = (m_i, 0)\), the
next intersection with the \(q_u\) axis will be at
\(x(t_{i+1}) = (m_{i+1},0)\) with \(|m_{i+1}| > |m_i|\).

@@

Setting \(I < 0\), the Poincar\'{e} section \(P_\mathcal{O}(\mu_0)\) is now
decreasing instead of increasing,
which means the constraint is dissipating energy.
